Knowledge \& the web: exercise session 6

### 2015-11-22 exercise_6/ex6.tex
conclusion

\documentclass[12pt,a4paper]{article}
% \usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{svg}
\usepackage{float}
\usepackage{url}
\usepackage{hyperref}

@@

\par The most obvious observation about this initial model is that female passengers with third class tickets are not yet classified very accurately, yet the tree did not split on it. In order to improve on this, we can adjust the splitting criterion to \emph{accuracy}. It will split nodes such that the accuracy of the tree increases. This results in a very sparse and large tree, which is not shown for exactly that reason. By changing the maximal depth of the tree, we can prevent it from getting too large. A maximal depth of 4 was found to provide good results. The result is shown in Figure~\ref{fig:tree_accuracy}.
\begin{figure}[htbp]
  \centering
  \includegraphics[width = .6\textwidth]{tree_accuracy}
  \captionsetup{width=.8\textwidth}
  \caption{Tree learned with \emph{accuracy} splitting criterion and maximum depth of 4.}
  \label{fig:tree_accuracy}
\end{figure}
As we can see, this achieved the desired result of splitting the undecided node in four new nodes, three of which perform clearly better. The final step is improving prediction for males. While its accuracy is quite high, the recall for surviving males is zero. This probably happens because the gain in criterion is not sufficiently high for the pre-pruning step to split the node. By lowering the minimum gain during pre-pruning, we enable the algorithm to split easier on nodes. Because the \texttt{Fare} node is already split until max depth, we expect that only the rightmost node (males) will be further split by this setting. Figure~\ref{fig:tree_final} confirms this assumption and shows the final model. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width = .8\textwidth]{tree_final}
  \captionsetup{width=.8\textwidth}
  \caption{Tree learned with \emph{accuracy} splitting criterion, maximum depth of 4 and minimum information gain of 0.01.}
  \label{fig:tree_final}
\end{figure}
\par This model is pretty intuitive. For females, Class is the most important deciding factor. Amongst the lowest class women, the ones that payed most for their ticket surprisingly have the lowest survival rate. For males, Age is the most important deciding factor. Children have a far higher survival rate than men.
\pagebreak
\section{Rule learning vs. Decision trees}

@@

\emph{Precision} is the fraction of retrieved instances that are relevant:
$$precision = \frac{\#\text{true alive}}{\#\text{true alive} + \#\text{false alive}}$$
\\
\\
In following table we give the accuracy, precision and recall of our best configuration.
From these scores we decided what was in our opinion the best method.

>>>>

\documentclass[11pt,a4paper]{article}
% \usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{svg}
\usepackage{float}
\usepackage{url}
\usepackage{hyperref}

@@

\par The most obvious observation about this initial model is that female passengers with third class tickets are not yet classified very accurately, yet the tree did not split on it. In order to improve on this, we can adjust the splitting criterion to \emph{accuracy}. It will split nodes such that the accuracy of the tree increases. This results in a very sparse and large tree, which is not shown for exactly that reason. By changing the maximal depth of the tree, we can prevent it from getting too large. A maximal depth of 4 was found to provide good results. The result is shown in Figure~\ref{fig:tree_accuracy}.
\begin{figure}[htbp]
  \centering
  \includegraphics[width = .6\textwidth]{tree_accuracy}
  \captionsetup{width=.8\textwidth}
  \caption{Tree learned with \emph{accuracy} splitting criterion and maximum depth of 4.}
  \label{fig:tree_accuracy}
\end{figure}
As we can see, this achieved the desired result of splitting the undecided node in four new nodes, three of which perform clearly better. The final step is improving prediction for males. While its accuracy is quite high, the recall for surviving males is zero. This probably happens because the gain in criterion is not sufficiently high for the pre-pruning step to split the node. By lowering the minimum gain during pre-pruning, we enable the algorithm to split easier on nodes. Because the \texttt{Fare} node is already split until max depth, we expect that only the rightmost node (males) will be further split by this setting. Figure~\ref{fig:tree_final} confirms this assumption and shows the final model. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width = .8\textwidth]{tree_final}
  \captionsetup{width=.8\textwidth}
  \caption{Tree learned with \emph{accuracy} splitting criterion, maximum depth of 4 and minimum information gain of 0.01.}
  \label{fig:tree_final}
\end{figure}
\par This model is pretty intuitive. For females, Class is the most important deciding factor. Amongst the lowest class women, the ones that payed most for their ticket surprisingly have the lowest survival rate. For males, Age is the most important deciding factor. Children have a far higher survival rate than men.

\section{Rule learning vs. Decision trees}

@@

\emph{Precision} is the fraction of retrieved instances that are relevant:
$$precision = \frac{\#\text{true alive}}{\#\text{true alive} + \#\text{false alive}}$$

Finally \emph{recall} (or sensitivity) is the fraction of relevant instances that are retrieved:
$$recall = \frac{\#\text{true alive}}{\#\text{true alive} + \#\text{false dead}}$$
\\
\\
In following table we give the accuracy, precision and recall of our best configuration from both rule learning and decision trees. With best we mean the highest \textbf{accuracy}, \textbf{precision} and \textbf{recall}.
\begin{figure}[H]
\centering
\begin{tabular}{c|c|c|c}

• & Accuracy & Precision & Recall \\ 
\hline 
Rule learning & 78.78\% & 75.95\% & 66.39\% \\ 
\hline 
Decision trees & 81.46\% & 81.37\% & 71.26\% \\

\end{tabular} 
\caption{Our results}
\end{figure}
From these scores it is obvious to see that the clustering with a decision tree is a lot better. With our configurations decision trees outperforms rule learning in every performance measure.
